---
title: "Customer Satisfaction Paper"
output:
  html_document:
    df_print: paged
---

## Satisfaction data
I have downloaded the data from [hcahpsonline](http://www.hcahpsonline.org/en/summary-analyses/). The tables were in PDF format, as in this [example](http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/results/2017-12_summary-analyses_state-results.pdf). In order to extract these tables, I used the function extract_areas from the package tabulizer, which essentially asks me to select the area where the table is and then it saves it as list object with a data frame per page in the document. 

The process is not perfect, but I have 36 files containing the quarterly data ranging from October 2006 to March 2017. I still need to clean it, though. 

```{r echo = FALSE, eval=FALSE}
library("tabulizer")

link1 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/march-2008-public-report-october-2006--june-2007-discharges.pdf"
link2 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/summer-2008-public-report-october-2006--september-2007-discharges.pdf"
link3 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/september-2008-public-report-january-2007--december-2007-discharges.pdf"
link4 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/december-2008-public-report-april-2007--march-2008-discharges.pdf"
link5 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/march-2009-public-report-july-2007--june-2008-discharges.pdf"
link6 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/june-2009-public-report-october-2007--september-2008-discharges.pdf"
link7 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/september-2009-public-report-january-2008--december-2008-discharges.pdf"
link8 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/december-2009-public-report-april-2008--march-2009-discharges.pdf"
link9 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/march-2010-public-report-july-2008--june-2009-discharges.pdf"
link10 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/july-2010-public-report-october-2008--september-2009-discharges.pdf"
link11 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/september-2010-public-report-january-2009--december-2009-discharges.pdf"
link12 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/december-2010-public-report-april-2009--march-2010-discharges.pdf"
link13 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/april-2011-public-report-july-2009--june-2010-discharges.pdf"
link14 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/summer-2011-public-report-october-2009--september-2010-discharges.pdf"
link15 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/october-2011-public-report-january-2010--december-2010-discharges.pdf"
link16 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/january-2012-public-report-april-2010---march-2011-discharges.pdf"
link17 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/spring-2012-public-report-july-2010---june-2011-discharges.pdf"
link18 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/july-2012-public-report-october-2010---september-2011-discharges.pdf"
link19 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/october-2012-public-report-january-2011---december-2011-discharges.pdf"
link20 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/december-2012-public-report-april-2011---march-2012-discharges.pdf"
link21 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/april-2013-public-report-july-2011---june-2012-discharges.pdf"
link22 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/july-2013-public-report-october-2011---september-2012-discharges.pdf"
link23 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/october-december-2013-public-report-january-2012---december-2012-discharges.pdf"
link24 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/december-2013-january-2014-public-report-april-2012---march-2013-discharges.pdf"
link25 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/april-2015-public-report-july-2013-june-2014-discharges.pdf"
link26 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/july-2015-public-report-october-2013-to-september-2014-discharges.pdf"
link27 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/october-2015-public-report-january-2014-to-december-2014-discharges.pdf"
link28 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/december-2015-public-report-april-2014-to-march-2015-discharges.pdf"
link29 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/april-may-2016-public-report-july-2014---june-2015-discharges.pdf"
link30 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/july-2016-public-report-october-2014---september-2015-discharges.pdf"
link31 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/october-2016-public-report-january-2015---december-2015-discharges.pdf"
link32 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/december-2016-public-report-april-2015---march-2016-discharges.pdf"
link33 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/april-2017-public-report-july-2015---june-2016-discharges.pdf"
link34 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/summary-results/july-2017-public-report-october-2015--september-2016-discharges.pdf"
link35 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/results/2017-10_summary-analyses_states-results.pdf"
link36 <- "http://www.hcahpsonline.org/globalassets/hcahps/summary-analyses/results/2017-12_summary-analyses_state-results.pdf"
download.file(link1, "file1.pdf",mode = "wb")
download.file(link2, "file2.pdf",mode = "wb")
download.file(link3, "file3.pdf",mode = "wb")
download.file(link4, "file4.pdf",mode = "wb")
download.file(link5, "file5.pdf",mode = "wb")
download.file(link6, "file6.pdf",mode = "wb")
download.file(link7, "file7.pdf",mode = "wb")
download.file(link8, "file8.pdf",mode = "wb")
download.file(link9, "file9.pdf",mode = "wb")
download.file(link10, "file10.pdf",mode = "wb")
download.file(link11, "file11.pdf",mode = "wb")
download.file(link12, "file12.pdf",mode = "wb")
download.file(link13, "file13.pdf",mode = "wb")
download.file(link14, "file14.pdf",mode = "wb")
download.file(link15, "file15.pdf",mode = "wb")
download.file(link16, "file16.pdf",mode = "wb")
download.file(link17, "file17.pdf",mode = "wb")
download.file(link18, "file18.pdf",mode = "wb")
download.file(link19, "file19.pdf",mode = "wb")
download.file(link20, "file20.pdf",mode = "wb")
download.file(link21, "file21.pdf",mode = "wb")
download.file(link22, "file22.pdf",mode = "wb")
download.file(link23, "file23.pdf",mode = "wb")
download.file(link24, "file24.pdf",mode = "wb")
download.file(link25, "file25.pdf",mode = "wb")
download.file(link26, "file26.pdf",mode = "wb")
download.file(link27, "file27.pdf",mode = "wb")
download.file(link28, "file28.pdf",mode = "wb")
download.file(link29, "file29.pdf",mode = "wb")
download.file(link30, "file30.pdf",mode = "wb")
download.file(link31, "file31.pdf",mode = "wb")
download.file(link32, "file32.pdf",mode = "wb")
download.file(link33, "file33.pdf",mode = "wb")
download.file(link34, "file34.pdf",mode = "wb")
download.file(link35, "file35.pdf",mode = "wb")
download.file(link36, "file36.pdf",mode = "wb")

file1 <- extract_areas("file1.pdf", method = "data.frame")
file2 <- extract_areas("file2.pdf", method = "data.frame")
file3 <- extract_areas("file3.pdf", method = "data.frame")
file4 <- extract_areas("file4.pdf", method = "data.frame")
file5 <- extract_areas("file5.pdf", method = "data.frame")
file6 <- extract_areas("file6.pdf", method = "data.frame")
file7 <- extract_areas("file7.pdf", method = "data.frame")
file8 <- extract_areas("file8.pdf", method = "data.frame")
file9 <- extract_areas("file9.pdf", method = "data.frame")
file10 <- extract_areas("file10.pdf", method = "data.frame")
file11 <- extract_areas("file11.pdf", method = "data.frame")
file12 <- extract_areas("file12.pdf", method = "data.frame")
file13 <- extract_areas("file13.pdf", method = "data.frame")
file14 <- extract_areas("file14.pdf", method = "data.frame")
file15 <- extract_areas("file15.pdf", method = "data.frame")
file16 <- extract_areas("file16.pdf", method = "data.frame")
file17 <- extract_areas("file17.pdf", method = "data.frame")
file18 <- extract_areas("file18.pdf", method = "data.frame")
file19 <- extract_areas("file19.pdf", method = "data.frame")
file20 <- extract_areas("file20.pdf", method = "data.frame")
file21 <- extract_areas("file21.pdf", method = "data.frame")
file22 <- extract_areas("file22.pdf", method = "data.frame")
file23 <- extract_areas("file23.pdf", method = "data.frame")
file24 <- extract_areas("file24.pdf", method = "data.frame")
file25 <- extract_areas("file25.pdf", method = "data.frame")
file26 <- extract_areas("file26.pdf", method = "data.frame")
file27 <- extract_areas("file27.pdf", method = "data.frame")
file28 <- extract_areas("file28.pdf", method = "data.frame")
file29 <- extract_areas("file29.pdf", method = "data.frame")
file30 <- extract_areas("file30.pdf", method = "data.frame")
file31 <- extract_areas("file31.pdf", method = "data.frame")
file32 <- extract_areas("file32.pdf", method = "data.frame")
file33 <- extract_areas("file33.pdf", method = "data.frame")
file34 <- extract_areas("file34.pdf", method = "data.frame")
file35 <- extract_areas("file35.pdf", method = "data.frame")
file36 <- extract_areas("file36.pdf", method = "data.frame")

df1 <- rbind( as.data.frame(file1[1]),as.data.frame(file1[2]))
df2 <- rbind( as.data.frame(file2[1]),as.data.frame(file2[2]))
df3 <- rbind( as.data.frame(file3[1]),as.data.frame(file3[2]))
df4 <- rbindlist(list(as.data.frame(file4[1]),as.data.frame(file4[2])),use.names = FALSE)
df5 <- rbind( as.data.frame(file5[1]),as.data.frame(file5[2])) #
df6 <- rbind( as.data.frame(file6[1]),as.data.frame(file6[2]))
df7 <- rbind( as.data.frame(file7[1]),as.data.frame(file7[2])) #
df8 <- rbind( as.data.frame(file8[1]),as.data.frame(file8[2])) #
df9 <- rbind( as.data.frame(file9[1]),as.data.frame(file9[2])) #
df10 <- rbind( as.data.frame(file10[1]),as.data.frame(file10[2])) #
df11 <- rbind( as.data.frame(file11[1]),as.data.frame(file11[2]))
df12 <- rbind( as.data.frame(file12[1]),as.data.frame(file12[2]))
df13 <- rbind( as.data.frame(file13[1]),as.data.frame(file13[2]))
df14 <- rbind( as.data.frame(file14[1]),as.data.frame(file14[2]))
df15 <- rbind( as.data.frame(file15[1]),as.data.frame(file15[2]))
df16 <- rbind( as.data.frame(file16[1]),as.data.frame(file16[2]))
df17 <- rbind( as.data.frame(file17[1]),as.data.frame(file17[2]))
df18 <- rbind( as.data.frame(file18[1]),as.data.frame(file18[2]))
df19 <- rbind( as.data.frame(file19[1]),as.data.frame(file19[2]))
df20 <- rbind( as.data.frame(file20[1]),as.data.frame(file20[2]))
df21 <- rbind( as.data.frame(file21[1]),as.data.frame(file21[2]))
df22 <- rbind( as.data.frame(file22[1]),as.data.frame(file22[2]))
df23 <- rbind( as.data.frame(file23[1]),as.data.frame(file23[2])) 
df24 <- rbind( as.data.frame(file24[1]),as.data.frame(file24[2]))
df25 <- rbind( as.data.frame(file25[1]),as.data.frame(file25[2]))
df26 <- rbind( as.data.frame(file26[1]),as.data.frame(file26[2]))
df27 <- rbind( as.data.frame(file27[1]),as.data.frame(file27[2]))
df28 <- rbind( as.data.frame(file28[1]),as.data.frame(file28[2]))
df29 <- rbind( as.data.frame(file29[1]),as.data.frame(file29[2]))
df30 <- rbind( as.data.frame(file30[1]),as.data.frame(file30[2]))
df31 <- rbind( as.data.frame(file31[1]),as.data.frame(file31[2]))
df32 <- rbind( as.data.frame(file32[1]),as.data.frame(file32[2]))
df33 <- rbind( as.data.frame(file33[1]),as.data.frame(file33[2]))
df34 <- rbind( as.data.frame(file34[1]),as.data.frame(file34[2]))
df35 <- rbind( as.data.frame(file35[1]),as.data.frame(file35[2]))
df36 <- rbind( as.data.frame(file36[1]),as.data.frame(file36[2]))
```

## Weather data
The weather data should start at least 2 years prior the beggining of the satisfaction data. That is, October 2004. I will start with January 2004. 

The data also has to be a the state level. This means that we need the data for the city's in the state that contain at least 80% of the population of the state. 

There is a package called "rwunderground" which allows for easy extraction of historical data from the website. It is free, however, Wunderground limits the number of "calls" to 10 per minute and 500 per day. This means that, to get the data we want for free, we will have to pay with time. A call appears to be a day, so that's a lot. 

Considering we need 13 years, for 50 states, each with at least 3 cities, it will take `r (13*365*50*3)/500/30` months to download this data. 

```{r eval = FALSE}
library(rwunderground)
library(XML)
library(RCurl)
library(rlist)
WApiKey <- "4226731577c71f1a"

data <- history_daily(set_location(territory = "New York", city = "New York"), date = "20040101",key = WApiKey)
data <- history_range(set_location(territory = "New York", city = "New York"), date_start = "200401", date_end = "200401", key = WApiKey)

link <- "https://www.wunderground.com/history/airport/KNYC/2004/1/1/CustomHistory.html?dayend=1&monthend=1&yearend=2005&req_city=&req_state=&req_statename=&reqdb.zip=&reqdb.magic=&reqdb.wmo="
readHTMLTable(getURL(link), which = 2)
```

